### 2016-11-21 intro.tex
Added James' comments

\section{The Cosmological Context of Galaxy Formation}
The modern picture of galaxy formation sits within the context of a $\Lambda$ +
Cold Dark Matter  ($\Lambda CDM$) cosmology
\citep{Rees1977,White1978,Blumenthal1984,Mo1998}.  This is a universe which has
a no curvature ($\Omega = 1$), and today has the majority of its mass-energy in
a cosmological constant $\Lambda$ ($\Omega_\Lambda \sim 0.7$), and most of its
matter in a collisionless form (dark matter) ($\Omega_m \sim 0.3$)
\citep{Planck2014}.  Roughly $5\%$ of the universe exists as baryons ($\Omega_b
\sim 0.05$), which existed as a nearly isothermal ($\delta T/T\sim10^{-5}$) gas
of hydrogen and helium at the earliest observational time.  All of this began
roughly $13.8$ billion years ago, with the universe emerging from a state of
infinite temperature and density, expanding to form what we see today.  The
primordial elemental abundances, with $\sim4/5$ of the universe's baryons in
hydrogen, and the remainder in helium, was established just a few minutes after
the big bang, during a brief period of ``Big Bang Nucleosynthesis''
\citep{Alpher1948}.  The remaining heavier elements have been formed via stellar
nucleosynthesis, through fusion in the cores of stars and in energetic processes
that take place during supernovae (SN) \citep{Wagoner1967}.

Much of what we know about the cosmology of our universe comes from observations
of the first light to escape the universe's surface of last scattering as it
became optically thin: the cosmic microwave background radiation (CMB).  Prior
to a redshift of $z\approx1100$, the universe had a high ionization fraction
such that the mean free path for photons was short, making all of space
effectively opaque.  Once the universe became cool enough that hydrogen became
fully neutral, the optical depth fell low enough that photons could free-stream,
decoupled from baryons.  This ultraviolet light we observe today, redshifted to
microwaves with a temperature of $\sim2.7\;\rm{K}$.  When the power spectrum of
this radiation is measured \citep{Spergel2003}, features in that spectrum can be
used to determine the components of the universe, as well as the spectrum of
density perturbations that will eventually form galaxies
\citep{Press1974,Peebles1980}.

The formation of galaxies within the universe began with the gravitational
collapse of small, linear density perturbations that were likely seeded by
quantum fluctuations amplified by inflation
\citep{Guth1981,Linde1982}.  These perturbations grow through
gravitational collapse.  Until $z\sim1100$, only dark matter perturbations were
able to grow.  Prior to this, baryons were coupled to photons, causing density
perturbations to simply oscillate as stable sound waves.  This early collapse
of matter meant that when decoupling occurred at $z\sim1100$, baryons were able
to collapse into the deeper potential wells grown out of dark-matter density
perturbations.  These overdensities expand slower that the surrounding medium,
as they experience a drag from their own self-gravity.  Eventually, these
regions stop expanding altogether and begin collapsing at the turnaround time
$t_{turn}$, and virialize at $2t_{turn}$ \citep{Peebles1980}. The first
galaxies likely began to form around $z\sim20-50$.

The perturbation spectrum that forms the structure in which galaxies begins to
form is essentially a Gaussian random field, which \citet{Press1974} were able
to show produces a halo mass spectrum at a given time with a relatively simple form:
\begin{equation}
    \frac{dn}{dM}(M,t) =
    \sqrt{\frac{2}{\pi}}\frac{\rho_0\delta_c(t)}{M^2\sigma(M)} 
    \left\lvert\frac{d\ln{\sigma}}{d\ln{M}}\right\rvert
    \exp{\left[-\frac{\delta^2_c(t)}{2\sigma^2(M)}\right]}
\end{equation}
In this equation, $\rho_0$ is the mean density of the universe, and
$\delta_c(t)$ is the critical overdensity for collapse at a given time $t$.  The
mass variance spectrum, $\sigma^2(M)$ can be determined observationally using
the CMB power spectrum \citep{Fixsen1996,Spergel2003,Planck2015}.  This function
gives a distribution of masses for the dark matter halos that galaxies will
begin to form within.  As these halos collapse, gravitational tugs from their
surroundings will impart torques on them \citep{Barnes1987}, causing halos to
have a distribution of angular momenta.  High resolution N-body simulations have
found that dark matter halos have both a universal density profile
\citep{Navarro1996,Merritt2006} as well as a universal angular momentum profile
\citep{Bullock2001}.  These dark matter halos all serve as the environment in
which the collapse of baryons and formation of stars take place.

The collapse of baryons within halos will depend strongly on whether the
post-virialized gas can cool radiatively \citep{Rees1977}.  As gas accretes onto
a halo with overdensity $\Delta$ it passes through a shock that raises its
temperature to the virial temperature:
\begin{equation}
    T_{virial} = \frac{2}{3}\frac{G\mu m_H}{k_B}\left({\frac{4\pi\rho_0\Delta
    M^2}{3}}\right)^{1/3}
\end{equation}
This means that more massive halos have larger virial temperatures, and,
depending on the mode of radiative cooling (atomic cooling from primordial
hydrogen, molecular hydrogen cooling, or cooling through metal lines), this may
give cooling times shorter than the current age of the universe.  These halos 
are the ones in which galaxy formation may begin, as their hot gaseous halos may
can collapse to form a disc where star formation can occur.  In some halos,
these virial shocks may be unstable to cooling, allowing the direct collapse of
cold gas onto the disc \citep{Birnboim2003}.

As gas does begin to cool within these halos, it will lose pressure support (by
radiating away thermal energy), but {\it not} angular momentum.  This means that
the baryons, which typically carry the same specific angular momentum as the
halo itself, will flatten to form an exponential disc as they cool
\citep{Fall1980,Mo1998}.  Streams of gas accreting along the cosmic web may also
provide angular momentum to the forming disc \citep{Dekel2006}.  Within this
disc, an interstellar medium (ISM) will form.  This ISM is multiphase,
containing both hot ($T\sim10^5\;\rm{K}$) and cold ($T\sim10-100\;\rm{K}$) in
quasistatic equilibrium \citep{McKee1977}.
\begin{figure}
    \includegraphics[width=0.7\textwidth]{mass_function.eps}
    \caption[Galaxy mass function]{The observed mass function for galaxies in
    the nearby universe (solid line) does not match the mass function for dark
    matter halos given in equation 1, scaled by the baryon fraction (dashed
    line).  This tells us that there are additional physical processes involved
    in the formation of galaxies within their dark matter halos.  The most
    likely candidate for this is feedback from stars and black holes.  Of
    particular interest is the peak in star formation efficiency near
    $M_{200}=10^{12}M_\odot$.  \textit{Image credit: Adapted from figure 1 in
    \citet{Ferrero2012}}}.
\end{figure}

These are of course all highly idealized, simple models for the earliest stages
and broadest strokes of galaxy formation.  Each of these halos will begin to
undergo mergers and interactions \citet{Navarro1993,Kaufmann1999,Cole2000} with
each other that can strip stars and gas through tidal and ram pressure forces
\citep{Gunn1972,Dressler1980}.  Inside each of these halos, the process of disc
collapse, the formation of a multiphase ISM, star formation, and feedback from
those stars involves the complex, nonlinear interplay of radiation,
hydrodynamics, magnetic fields, and gravity on length scales as much as
$10^{11}$ times smaller than the virial radius.  None of this physics is simple
enough to be fully captured by these sorts of analytic models, and as figure 1.1
shows, these processes are important to ultimately establishing the population
of galaxies we see today.  Numerical simulations have therfore become a keystone for
galaxy formation research.

\section{Numerical Simulations of Galaxies}
Simulations of galaxies actually predate {\it numerical} simulations of
galaxies.  \citet{Holmberg1941} presented, as the author termed it, a ``new
integration procedure'' that was the first attempt to simulate the
evolution of galaxy.  By relying on the fact that light flux, just as gravity,
scales as $r^{-2}$, Holmberg was able to construct a simulated galaxy disc out
of light bulbs, and then using photocells to measure the flux at each bulb's
position, repositioned the bulbs using the forces ``simulated'' by the
light-measuring process.  This allowed him to painstakingly trace the tidal
interactions of passing galaxies as they fell into a galaxy cluster.

@@

The earliest numerical simulations of galaxy formation focused on the effects of
gravity alone.  A classic example of this are the simulations presented in
\citet{Toomre1972}, which showed that the irregular structures seen in galaxies
like NGC 4676 (``The Mice'') or NGC4038/4039 (``The Antennae Galaxies'') are the
results of major mergers between two disc galaxies, as is clear from figure 1.2.
These simulations used discs composed of a mere 120 test particles, and ignored
the effects of self-gravity between these particles.  Despite their crudeness,
these simulations definitively showed that galactic bridges and tails could be
formed by the tidal interactions of two passing/merging galaxies.  The IBM
360/95 that \citet{Toomre1972} (with an inflation-adjusted purchase price in the
millions of dollars) used for their simulations was capable of 5.5 million
instructions per second, roughly $0.2\%$ the speed of a 35\$ Raspberry Pi 3
available today.

@@

Gravity is typically solved using algorithms designed to minimize the workload
that would be needed for a direct summation of each particle or grid cell
($\mathcal{O}(N^2)$ for N grid cells/particles).  This becomes intractable for
larger numbers of resolution elements, as the workload will scale as the sixth
power of the linear resolution (an order of magnitude better linear resolution
will be $10^6$ times more expensive!).  In order to reduce these costs,
algorithms have been developed that allow for more efficient gravitational
force calculations.  Tree-based methods \citep{Barnes1986} reduce the
computational cost of long-ranged (and thus less significant) force
calculations by using space-partitioning trees.  Short-range force calculations
are done particle-by-particle, while long-range forces are calculated against
larger tree cells, thus reducing the cost of calculating gravity to
$\mathcal{O}(N\log N)$.  Particle-Particle/Particle-Mesh ($P^3M$) methods
\citep{Couchman1991} produce similar scalings, but rather than using a
spatial tree rely on a spectral Fourier-based approach to calculate long-ranged
interactions.  Recently, the Fast Multipole Method \citep{Greengard1987} has
been adapted for calculation of gravitational forces
\citep{Dehnen2002,Hahn2013} using a multipole expansion of Green's functions.
This allows for extremely efficient gravity calculation, scaling linearly with
particle number ($\mathcal{O}(N)$).

Hydrodynamical solvers usually come in one of two classes: Eulerian or
Lagrangian.  Eulerian methods discretize space, breaking a simulation volume
into a grid of cells that each contains scalar and vector quantities that are
advected according to the solver scheme.  Eulerian schemes typically allow for
higher-order solvers, lower intrinsic noise, and better resolved discontinuities
\citep{Teyssier2002,Stone2008,Bryan2014}.  They suffer from intrinsic mixing and
diffusion that scales with the flow velocity relative to the grid
\citep{Agertz2007,Tasker2008}.  Lagrangian methods instead discretize mass,
breaking the gaseous volume into discrete points, each of which is moved to
advect the quantities it contains, with smoothing functions used to interpolate
values in between particles; hence the name for the most common Lagrangian
method, Smoothed Particle Hydrodynamics (SPH).  Lagrangian methods couple easily
to tree codes for gravity calculation, are Galilean invariant, and intrinsically
conserve mass, energy, and momentum (linear and angular)
\citep{Katz1996,Wadsley2004,Springel2005}.  They require the use of artificial
viscosity to handle shocks, and as such have difficulty in capturing subsonic
turbulence \citep{Bauer2012}. Recently, new methods have begin to hybridize
features of each of these classes as well \citep{Springel2010,Hopkins2015}.

Modern galaxy simulations also now use a variety of strategies for judiciously
applying computational power where it is most needed.  Eulerian methods will
commonly use \textit{Adaptive Mesh Refinement} to increase resolution in regions
of interest (typically where gas is densest).  Lagrangian methods automatically
do this, as a denser region definitionally must contain more particles.  Beyond
the choice of algorithms, the choice of what type of simulation to set up in the
initial conditions is also important.  To study large populations of galaxies,
along with large-scale structure, large-volume cosmological boxes with volumes
$>>10^4\;\rm{Mpc^3}$, but with spatial resolutions $\sim \rm{kpc}$ are used.
Studying how individual galaxies form is typically done with cosmological zoom
simulations, where a comparatively cheap dark-matter only simulation is first
run to select a region of interest, which is then resimulated with higher
resolution and baryonic physics included \citep{Quinn1992}. Zoom-in simulations
can typically achieve an order of magnitude or better resolution than full
cosmological boxes.  Isolated galaxy disc simulations cannot show how galaxies
form in a cosmological environment, but can be run with $\sim \rm{pc}$
resolution to study closely how internal processes within a galaxy take place
\citep{Hopkins2011,Benincasa2015}.

The earliest attempt to track the formation of a galaxy in a cosmological
context with the effects of star formation and feedback included was by
\citet{Katz1992}.  This simulation used the Lagrangian TREESPH method
\citep{Hernquist1989} to evolve a parcel of gas and dark matter from an initial
linear perturbation, including small-scale power derived from CDM predictions
\citep{Zeldovich1970,Peebles1982}.
Gas was allowed to cool radiatively using rates calculated for primordial gas.
This study introduced a model for star formation that is still the standard for
most simulations of galaxy formation.  Observations \citep{Schmidt1959} have
suggested that the star formation rate is a power law of the gas density:
$\dot\rho_*\propto\rho_{gas}^\alpha$, with $\alpha=1-2$.  If it is assumed that
gas collapses to form stars in some multiple $c_*$ of its freefall time,
$t_{ff}$, a simple star formation model can be built with the form:
\begin{equation}
    \dot\rho_* = \frac{c_*\rho_{gas}}{t_{ff}} = c_*\sqrt{\frac{\rho^3_{gas}}{4\pi G}}
\end{equation}
\citet{Katz1992} used a $c_*$ parameter of $0.1$, which is within the range of
what is typically used in simulations today ($1-100\%$).  On top of the model
for star formation, feedback was included without the use of any significant
subgrid model:  $10^{51}\;\rm{erg/SN}$ worth of energy was injected back into
the ISM by smoothing it over an SPH kernel.  One of the key findings of
\citet{Katz1992} was that the vast majority of SN energy was radiated away,
making them very ineffective at regulating star formation within the galaxy.  In
the end, \citet{Katz1992} found that the peak starformation in the simulated galaxy was
nearly 10 times greater than in observed galaxies.  \citet{Thacker2000} later
showed that the smoothing of energy used by \citet{Katz1992} resulted in a large
overestimation of the cooling rates in feedback-heated gas. Since then,
much work has gone into developing models for SN feedback that address the issue
of overcooling that was present in these earlier simulations.

\section{The Importance of Feedback}
We now know that galaxy and star formation is not a one way process of
gravitational collapse.  A number of internal processes (``feedback'') within
galaxies can liberate energy greater than the gravitational binding energy of
molecular clouds (and indeed even comparable to the binding energy of the entire
halo).  Analytic calculations that omit feedback processes found that gas should
rapidly collapse into the smallest halos at high redshift, wildy overproducing
stars in small galaxies at high redshift (the ``cooling catastrophe'')
\citep{Cole2001,Benson2003}.  Simulations of galaxies like our own develop unrealistically
compact morphologies without feedback \citep{Stinson2006}.  Without feedback, simulations of
individual molecular clouds have star formation rates an order of magnitude
too high \citep{Agertz2013}.  Aside from regulating star formation and the
growth of galaxies, there are also observations of the large-scale impact of
feedback. Lyman $\alpha$ forest observations probing the intergalactic medium
(IGM) have found that the IGM is polluted with metal ions at great distances
from any galaxies \citep{Sargent1988,Songaila1996,Dave1998}.  These metals must have
been ejected by galactic outflows from the discs of galaxies where they formed.
As figure 1.3 shows, fast, massive outflows have been observed leaving the discs of
starburst galaxies and quasar host galaxies \citep{Veilleux2005,Werk2014}.
These outflows must be powered by feedback.

\begin{figure}
    \includegraphics[width=0.8\textwidth]{M82.eps}
	\caption[Massive outflows in M82]{The evidence for feedback is no clearer
	than here in the Cigar Galaxy, M82.  The Hubble Space Telescope reveals massive
	outflows of hot ionized gas through the red $H\alpha$ emission as gas is
	blasted out of the galaxy in a bipolar flow. \textit{Image credit: NASA, ESA
	and the Hubble Heritage Team STScI/AURA}}
\end{figure}

The two primary sources of feedback anergy are massive stars and supermassive
black holes (SMBHs).  Massive stars ($M>5M_\odot$) live short, violent lives.
During their lives, they will drive fast $v>1000\;\rm{km/s}$ stellar winds
\citep{Weaver1977}, and heat their surrounding gas with UV radiation
\citep{Krumholz2009}.  At the end of their lives, $4-30\;\rm{Myr}$ later, these
stars die in extremely energetic core-collapse SN, liberating
$\sim10^{51}\;\rm{erg}$ \citep{Leitherer1999}.  Figure 1.4 shows the total
budget of available energy liberated by a population of stars.  This energy
imparts heat and momentum to the surrounding ISM, and can accelerate cosmic rays
in the shocks that result \citep{Bell1978}.  In galaxies with actively growing
SMBHs, the accretion disc that forms as gas funnels into the black hole can
reach temperatures exceeding $10^9\;\rm{K}$ through viscous heating
\citep{Antonucci1993}.  This makes
these accretion discs some of the most luminous objects in the universe, and can
heat the gas that surrounds the galaxy to choke off the accretion of new
material to form stars.

>>>>

\section{The Cosmological Context of Galaxy Formation}
The modern picture of galaxy formation sits within the context of a $\Lambda$ 
Cold Dark Matter  ($\Lambda CDM$) cosmology
\citep{Rees1977,White1978,Blumenthal1984,Mo1998}.  This is a universe which has
a no curvature ($\Omega = 1$), and today has the majority of its mass-energy in
a cosmological constant $\Lambda$ ($\Omega_\Lambda \sim 0.7$), and most of its
matter in a collisionless form (dark matter) ($\Omega_m \sim 0.3$)
\citep{Planck2014}.  Roughly $5\%$ of the matter-energy density of the universe
exists as baryons ($\Omega_b \sim 0.05$), which existed as a nearly isothermal
($\delta T/T\sim10^{-5}$) gas of hydrogen and helium at the earliest
observational time.  All of this began roughly $13.8$ billion years ago, with
the universe emerging from a state of high temperature and density,
expanding to form what we see today.  The primordial elemental abundances, with
$\sim4/5$ of the universe's baryons in hydrogen, and the remainder in helium,
was established just a few minutes after the big bang, during a brief period of
``Big Bang Nucleosynthesis'' \citep{Alpher1948}.  The remaining heavier elements
have been formed via stellar nucleosynthesis, through fusion in the cores of
stars and in energetic processes that take place during supernovae (SN)
\citep{Wagoner1967}, or through interactions between cosmic rays and nucleii in
the interstellar medium (ISM) \citep{Reeves1970}.

Much of what we know about the cosmology of our universe comes from
observations of the first light to escape the universe's surface of last
scattering as it became optically thin: the cosmic microwave background
radiation (CMB).  Prior to a redshift of $z\approx1100$, the universe had a
high ionization fraction such that the mean free path for photons was short,
making all of space effectively opaque.  Once the universe became cool enough
that hydrogen became fully neutral, the optical depth fell low enough that
photons could free-stream, decoupled from baryons.  This ultraviolet light we
observe today, redshifted to microwaves with a temperature of
$\sim2.7\;\rm{K}$.  When the power spectrum of this radiation is measured
\citep{Spergel2003}, features in that spectrum can be used to determine the
components of the universe (with some ambiguity that can be resolved via SN Ia
and large scale structure measurements
\citealt{Riess1998,Perlmutter1999,Beutler2011,Blake2011}), as well as the
spectrum of density perturbations that will eventually form galaxies
\citep{Press1974,Peebles1980}.

The formation of galaxies within the universe began with the gravitational
collapse of small, linear density perturbations that were likely seeded by
quantum fluctuations amplified by inflation
\citep{Guth1981,Linde1982}.  These perturbations grow through
gravitational collapse.  Until $z\sim1100$, only dark matter perturbations were
able to grow.  Prior to this, baryons were coupled to photons, causing density
perturbations to oscillate as stable sound waves.  This early collapse
of matter meant that when decoupling occurred at $z\sim1100$, baryons were able
to collapse into the deeper potential wells grown out of dark-matter density
perturbations.  These overdensities expand slower that the surrounding medium,
as they experience a drag from their own self-gravity.  Eventually, these
regions stop expanding altogether and begin collapsing at the turnaround time
$t_{turn}$, and virialize at $2\;t_{turn}$ \citep{Peebles1980}. The first
galaxies likely began to form around $z\sim20-50$.

The perturbation spectrum that forms the structure in which galaxies begins to
form is essentially a Gaussian random field, which \citet{Press1974} were able
to show produces a halo mass spectrum at a given time with a form that matched
the matched the functional form for the observed galaxy luminosity function
\citet{Schechter1976}:
\begin{equation}
    \frac{dn}{dM}(M,t) =
    \sqrt{\frac{2}{\pi}}\frac{\rho_0\delta_c(t)}{M^2\sigma(M)} 
    \left\lvert\frac{d\ln{\sigma}}{d\ln{M}}\right\rvert
    \exp{\left[-\frac{\delta^2_c(t)}{2\sigma^2(M)}\right]}
\end{equation}
In this equation, $\rho_0$ is the mean density of the universe, and
$\delta_c(t)$ is the critical overdensity for collapse at a given time $t$.
The mass variance spectrum, $\sigma^2(M)$ can be determined observationally
using the CMB power spectrum \citep{Fixsen1996,Spergel2003,Planck2015} and an
assumption of a specific cosmology.  This function gives a distribution of
masses for the dark matter halos that galaxies will begin to form within.  As
these halos collapse, gravitational tugs from their surroundings will impart
torques on them \citep{Barnes1987}, causing halos to have a distribution of
angular momenta.  High resolution N-body simulations have found that dark
matter halos have both a universal density profile
\citep{Navarro1996,Merritt2006} as well as a universal angular momentum profile
\citep{Bullock2001}.  These dark matter halos all serve as the environment in
which the collapse of baryons and formation of stars take place.

The collapse of baryons within halos will depend strongly on whether the
post-virialized gas can cool radiatively \citep{Rees1977}.  As gas accretes onto
a halo with overdensity $\Delta$ it passes through a shock that raises its
temperature to the virial temperature:
\begin{equation}
    T_{virial} = \frac{2}{3}\frac{G\mu m_H}{k_B}\left({\frac{4\pi\rho_0\Delta
    M^2}{3}}\right)^{1/3}
\end{equation}
This means that more massive halos have larger virial temperatures, and,
depending on the mode of radiative cooling (atomic cooling from primordial
hydrogen, molecular hydrogen cooling, or cooling through metal lines), this may
give cooling times shorter than the current age of the universe.  These halos 
are the ones in which galaxy formation may begin, as their hot gaseous halos may
can collapse to form a central concentration where star formation can occur.  In some halos,
these virial shocks may be unstable to cooling, allowing the direct collapse of
cold gas onto the disc \citep{Birnboim2003}.

As gas does begin to cool within these halos, it will lose pressure support (by
radiating away thermal energy), but {\it not} angular momentum.  For large
spiral galaxies, this means that the baryons, which typically carry the same
specific angular momentum as the halo itself, will flatten to form an
exponential disc as they cool \citep{Fall1980,Mo1998}.  Streams of gas
accreting along the cosmic web may also provide angular momentum to the forming
disc \citep{Dekel2006}.  Within this disc, an ISM will form.  This ISM is
multiphase, containing both hot ($T\sim10^4\;\rm{K}$) and cold
($T\sim10-100\;\rm{K}$) in quasistatic equilibrium \citep{McKee1977}.  This
multiphase state is highly dependent on both the ISM metallicity and the local
heating rates (which themselves can vary with redshift)
\citep{Field1969,Wolfire1995,Norman1997}.
\begin{figure}
    \includegraphics[width=0.7\textwidth]{mass_function.eps}
    \caption[Galaxy mass function]{The observed mass function for galaxies in
    the nearby universe (solid line) does not match the mass function for dark
    matter halos given in equation 1, scaled by the baryon fraction (dashed
    line).  This tells us that there are additional physical processes involved
    in the formation of galaxies within their dark matter halos.  The most
    likely candidate for this is feedback from stars and black holes.  Of
    particular interest is the peak in star formation efficiency near
    $M_{200}=10^{12}M_\odot$.  \textit{Image credit: Adapted from figure 1 in
    \citet{Ferrero2012}}}.
\end{figure}

These are of course all highly idealized, simple models for the broadest strokes
of galaxy formation.  Each of these halos will begin to undergo mergers and
interactions \citet{Navarro1993,Kauffmann1999,Cole2000} with each other that can
strip stars and gas through tidal and ram pressure forces
\citep{Gunn1972,Dressler1980}.  Inside each of these halos, the process of disc
collapse, the formation of a multiphase ISM, star formation, and feedback from
those stars involves the complex, nonlinear interplay of radiation,
hydrodynamics, magnetic fields, and gravity on length scales as much as
$10^{11}$ times smaller than the virial radius.  None of this physics is simple
enough to be fully captured by analytic models. As figure 1.1 shows, the
observed mass function for galaxies does not match what would be expected from
total baryon budget for each halo, implying that at least some of these
processes must change the galactic star formation efficiency as a function of
halo mass, establishing the population of galaxies we see today.  Numerical
simulations have therfore become a keystone for galaxy formation research.

\section{Numerical Simulations of Galaxies}
N-body Simulations of galaxies actually predate {\it digital} simulations of
galaxies.  \citet{Holmberg1941} presented, as the author termed it, a ``new
integration procedure'' that was the first attempt to simulate the
evolution of galaxy.  By relying on the fact that light flux, just as gravity,
scales as $r^{-2}$, Holmberg was able to construct a simulated galaxy disc out
of light bulbs, and then using photocells to measure the flux at each bulb's
position, repositioned the bulbs using the forces ``simulated'' by the
light-measuring process.  This allowed him to painstakingly trace the tidal
interactions of passing galaxies as merged with each other.

@@

The earliest numerical simulations of galaxy formation focused on the effects of
gravity alone.  A classic example of this are the simulations presented in
\citet{Toomre1972}, which showed that the irregular structures seen in galaxies
like NGC 4676 (``The Mice'') or NGC4038/4039 (``The Antennae Galaxies'') are the
results of major mergers between two disc galaxies, as is clear from figure 1.2.
These simulations used discs composed of a mere 120 test particles, and ignored
the effects of self-gravity between these particles.  Despite their crudeness,
these simulations definitively showed that the morphologies of galactic bridges and tails could be
formed by the tidal interactions of two passing/merging galaxies.  The IBM
360/95 that \citet{Toomre1972} (with an inflation-adjusted purchase price in the
millions of dollars) used for their simulations was capable of 5.5 million
instructions per second, roughly $0.2\%$ the speed of a 35\$ Raspberry Pi 3
available today.

@@

Gravity is typically solved using algorithms designed to minimize the workload
that would be needed for a direct summation of each particle or grid cell
($\mathcal{O}(N^2)$ for N grid cells/particles).  This becomes intractable for
larger numbers of resolution elements, as the workload will scale as the sixth
power of the linear resolution (an order of magnitude better linear resolution
will be $10^6$ times more expensive!).  In order to reduce these costs,
algorithms have been developed that allow for more efficient gravitational
force calculations.  Tree-based methods \citep{Barnes1986} reduce the
computational cost of long-ranged (and thus less significant) force
calculations by using space-partitioning trees.  Short-range force calculations
are done particle-by-particle, while long-range forces are calculated against
larger tree cells, thus reducing the cost of calculating gravity to
$\mathcal{O}(N\log N)$.  Particle-Particle/Particle-Mesh ($P^3M$) methods
\citep{Couchman1991} produce similar scalings, but rather than using a
spatial tree rely on a spectral Fourier-based approach to calculate long-ranged
interactions.  Recently, the Fast Multipole Method \citep{Greengard1987} has
been applied to the calculation of gravitational forces
\citep{Dehnen2002,Hahn2013} using a multipole expansion of Green's functions.
This allows for extremely efficient gravity calculation, scaling linearly with
particle number ($\mathcal{O}(N)$).

Hydrodynamical solvers usually come in one of two classes: Eulerian or
Lagrangian.  Eulerian methods discretize space into fixed volume elements with
static boundaries.  By breaking a simulation volume into a grid of cells that
each contains scalar and vector quantities those quantities can be advected
between cells according to the solver scheme.  Eulerian schemes typically allow
for higher-order solvers, lower intrinsic noise, and better resolved
discontinuities \citep{Teyssier2002,Stone2008,Bryan2014}.  They suffer from
intrinsic mixing and diffusion that scales with the flow velocity relative to
the grid, poor integration of circular orbits, and costly timestep limits
\citep{Agertz2007,Tasker2008}.  Lagrangian methods instead discretize mass,
breaking the gaseous volume into discrete points, each of which moves at the
fluid velocity to advect the quantities it contains, with smoothing functions
used to interpolate values in between particles; hence the name for the most
common Lagrangian method, Smoothed Particle Hydrodynamics (SPH).  Lagrangian
methods couple easily to tree codes for gravity calculation, are Galilean
invariant, and intrinsically conserve mass, energy, and momentum (linear and
angular) \citep{Katz1996,Wadsley2004,Springel2005}.  SPH requires the use of
artificial viscosity to handle shocks, and as such have difficulty in capturing
subsonic turbulence \citep{Bauer2012}. Recently, new methods have begin to
hybridize features of each of these classes as well
\citep{Springel2010,Hopkins2015}.

Modern galaxy simulations also now use a variety of strategies for judiciously
applying computational power where it is most needed.  Eulerian methods will
commonly use \textit{Adaptive Mesh Refinement} to increase resolution in regions
of interest (typically where gas is densest).  Lagrangian methods automatically
do this, as a denser region definitionally must contain more fluid elements.

Beyond the choice of algorithms, the choice of what type of simulation to set
up in the initial conditions is also important.  To study large populations of
galaxies, along with large-scale structure, large-volume cosmological boxes
with volumes $>10^6;\rm{Mpc^3}$, but with spatial resolutions $\sim \rm{kpc}$
are used.  Studying how individual galaxies form is typically done with
cosmological zoom simulations, where a comparatively cheap dark-matter only
simulation is first run to select a region of interest, which is then
resimulated with higher resolution and baryonic physics included
\citep{Navarro1993}. Zoom-in simulations can typically achieve an order of
magnitude or better resolution than full cosmological boxes.  Isolated galaxy
disc simulations cannot show how galaxies form in a cosmological environment,
but can be run with $\sim \rm{pc}$ resolution to study closely how internal
processes within a galaxy take place \citep{Hopkins2011,Benincasa2016}.  
These details of these internal processes are discussed in section 1.3.

The earliest attempt to track the formation of a galaxy in a cosmological
context with the effects of star formation and feedback included was by
\citet{Katz1992}.  This simulation used the Lagrangian TREESPH method
\citep{Hernquist1989} to evolve a parcel of gas and dark matter from an initial
linear perturbation, including small-scale power derived from CDM predictions
\citep{Zeldovich1970,Peebles1982}.
Gas was allowed to cool radiatively using rates calculated for primordial gas.
This study introduced a model for star formation that is still the standard for
most simulations of galaxy formation.  Observations \citep{Kennicutt1998} have
suggested that the star formation rate is a power law of the gas density:
$\dot\rho_*\propto\rho_{gas}^\alpha$, with $\alpha=1-2$.  If it is assumed that
gas collapses to form stars in some multiple $c_*$ of its freefall time
\citep{Schmidt1959},
$t_{ff}$, a simple star formation model can be built with the form:
\begin{equation}
    \dot\rho_* = \frac{c_*\rho_{gas}}{t_{ff}} = c_*\sqrt{\frac{\rho^3_{gas}}{4\pi G}}
\end{equation}
\citet{Katz1992} used a $c_*$ parameter of $0.1$, which is within the range of
what is typically used in simulations today ($1-100\%$).  On top of the model
for star formation, feedback was included without the use of any significant
subgrid model:  $10^{51}\;\rm{erg/SN}$ worth of energy was injected back into
the ISM by smoothing it over a number of resolution elements.  One of the key
findings of \citet{Katz1992} was that the vast majority of SN energy was
radiated away, making them very ineffective at regulating star formation within
the galaxy.  In the end, \citet{Katz1992} found that the peak starformation in
the simulated galaxy was nearly 10 times greater than in observed galaxies.
\citet{Thacker2000} later showed that the smoothing of energy used by
\citet{Katz1992} resulted in a large overestimation of the cooling rates in
feedback-heated gas. Since then, much work has gone into developing models for
SN feedback that address the issue of overcooling that was present in these
earlier simulations.

\section{The Importance of Feedback}
\begin{figure}
    \includegraphics[width=\textwidth]{M82.eps}
	\caption[Massive outflows in M82]{The evidence for feedback is no clearer
	than here in the Cigar Galaxy, M82.  The Hubble Space Telescope reveals massive
	outflows of hot ionized gas through the red $H\alpha$ emission as gas is
	blasted out of the galaxy in a bipolar flow. \textit{Image credit: NASA, ESA
	and the Hubble Heritage Team STScI/AURA}}
\end{figure}
We now have strong evidence that galaxy and star formation is not a one way process of
gravitational collapse.  A number of internal processes (``feedback'') within
galaxies can liberate energy greater than the gravitational binding energy of
molecular clouds (and indeed even comparable to the binding energy of the entire
halo).  Analytic calculations that omit feedback processes found that gas should
rapidly collapse into the smallest halos at high redshift, wildy overproducing
stars in small galaxies at high redshift (the ``cooling catastrophe'')
\citep{Cole2001,Benson2003}.  Simulations of galaxies like our own develop unrealistically
compact morphologies without feedback \citep{Stinson2006}.  Without feedback, simulations of
individual molecular clouds have star formation rates an order of magnitude
too high \citep{Agertz2013}.  Aside from regulating star formation and the
growth of galaxies, there are also observations of the large-scale impact of
feedback. Lyman $\alpha$ forest observations probing the intergalactic medium
(IGM) have found that the IGM is polluted with metal ions at great distances
from any galaxies \citep{Sargent1988,Songaila1996,Dave1998}.  These metals must have
been ejected by galactic outflows from the ISM of galaxies where they formed.
As figure 1.3 shows, fast, massive outflows have been observed leaving the ISM of
starburst galaxies and quasar host galaxies \citep{Veilleux2005,Werk2014}.

The two primary sources of feedback energy are massive stars and supermassive
black holes (SMBHs).  Massive stars ($M>5M_\odot$) experience short, violent
lives.  During their lives, they will drive fast $v>1000\;\rm{km/s}$ stellar
winds \citep{Weaver1977}, and heat their surrounding gas with UV radiation
\citep{Krumholz2009}.  At the end of their lives, $4-30\;\rm{Myr}$ later, these
stars die in energetic core-collapse SN, liberating $\sim10^{51}\;\rm{erg}$
\citep{Wilson1985,Bethe1990}.  Figure 1.4 shows the total budget of available
energy liberated by a population of stars.  The energy released by a population
of stars with solar metallicity is $\sim 2\times10^{51}\;\rm{erg/M_\odot}$.
Most of this energy is released as photons, over the lifetime of each star.
These will couple poorly to the surrounding ISM, and so have little impact on
disrupting all but the densest gas.  While stellar winds typically have
comparable luminosity to supernovae, they occur for an order of magnitude less
time ($\sim4$ Myr vs. $\sim40$ Myr).  This means that supernovae, despite only
contributing $\sim1\%$ of the total energy released by a stellar population, may
have the greatest impact on galactic scales.  This energy imparts heat and
momentum to the surrounding ISM, and can accelerate cosmic rays in the shocks
that result \citep{Bell1978}.  In galaxies with actively growing SMBHs, the
accretion disc that forms as gas funnels into the black hole can reach
temperatures exceeding $10^9\;\rm{K}$ through viscous heating
\citep{Antonucci1993}.  This makes these accretion discs some of the most
luminous objects in the universe, and can heat the gas that surrounds the galaxy
to choke off the accretion of gas onto the galaxy, shutting off star formation.
